Independent discovery of a nulling pulsar with unusual sub-pulse drifting properties with the Murchison Widefield Array

### 2021-11-26 main.tex
Several changes to the text after getting lots of great feedback from co-authors.

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%%
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands.
%%   tighten      : Makes baselineskip slightly smaller, only works with
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much
%%                  longer author list but may be desirable for short
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\psr}{J0026$-$1955}
\newcommand{\psrdm}{20.869}
\newcommand{\psrdmerr}{0.005}
\newcommand{\psrnick}{J0036$-$1033}
\newcommand{\dmunit}{pc$\,$cm$^{-3}$}
\newcommand{\Sd}{S$_\rm{d}$}
\newcommand{\nullfraction}{77\%}
\newcommand{\ExB}{\ensuremath{{\bf E}\times{\bf B}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Pulsar with unusual sub-pulse drifting}
\shortauthors{McSweeney et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page
%% with this command:
%% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{figures/}}
%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

@@

\title{Discovery of a nulling pulsar with unusual sub-pulse
drifting properties with the Murchison Widefield Array}

@@

\author[0000-0002-0786-7307]{Samuel J. McSweeney}
\affiliation{ICRAR, Curtin University}

\author{Ramesh Bhat}
\affiliation{ICRAR, Curtin University}

\author{Nick Swainston}
\affiliation{ICRAR, Curtin University}

\author{Keegan Smith}
\affiliation{ICRAR, Curtin University}

\author{Sanjay Kudale}
%\affiliation{}

\author{Paul Hancock}
\affiliation{ICRAR, Curtin University}

\author{Willem van Straten}
%\affilitation{}

\author{Shi Dai}
%\affilitation{}

@@

\author{Steven J Tingay}
\affiliation{ICRAR, Curtin University}

\author{Melanie Johnston-Hollitt}
%\affilitation{}

\author{David Kaplan}
%\affilitation{}

\author{Mia Walker}
\affiliation{ICRAR, Curtin University}

@@

We report the discovery of PSR \psr{} with the Murchison Widefield Array (MWA), the second to be discovered in the ongoing Southern-sky MWA Rapid Two-meter (SMART) pulsar survey.
\psr{} has a period of $\sim1.306\,$s, a dispersion measure (DM) of $\sim\psrdm\,$\dmunit, and a nulling fraction of $\sim$ \nullfraction{}.
The detection of this pulsar highlights the advantages of the survey's long dwell times ($\sim80\,$min), which is sensitive to the expected population of bright, intermittent pulsars with long nulls.
A single-pulse analysis also reveals a complex sub-pulse drifting behavior, including both rapid changes of the drift rate characteristic of mode switching pulsars, as well as a slow, consistent evolution of the drift rate within modes. In some longer drift sequences, interruptions in the otherwise smooth drift rate evolution occur preferentially at a particular phase, typically lasting a few pulses.
These properties make this pulsar an ideal test bed for prevailing models of drifting behavior such as the carousel model.

@@

Pulsars, or rapidly rotating neutron stars, are the sites of some of the highest energy physical processes in the universe, owing to the intense gravitational and magnetic fields that surround them.
Discovered first via their lighthouse-like beam of radio emission \citep{Hewish1968}, which we detect as a series of regularly spaced pulses, it was quickly realised that the physics governing the relativistic plasma that generates this beam was not well understood---a state of affairs that persists to the present day.
The uniqueness of each pulsar's emission signature provides a wealth of data that can be used to test candidate emission mechanisms, and it has been proven again and again over the past several decades that in-depth studies of individual pulsars can be just as useful for uncovering vital clues for understanding the population as a whole.
This, in part, motivates efforts to find new pulsars via large scale pulsar surveys, which continue to be conducted up to the present day.

One of the most recent such surveys is currently under way at the Murchison Widefield Array \citep[MWA;][]{Tingay2013}, a latest-generation aperture array telescope located in Western Australia.
With its Voltage Capture System \citep[VCS;][]{Tremblay2015} and offline tied-array beamforming software \citep{Ord2019,McSweeney2020}, it has proven to be a powerful instrument for investigating a variety of pulsar phenomena such as single-pulse studies \citep{McSweeney2017}, spectral analysis \citep{Meyers2017}, and studies of interstellar medium (ISM) propagation effects and related timing applications \citep{Bhat2018,Kaur2019}.
With the advent of the Phase II upgrade of the MWA \citep{Wayth2018} in its compact configuration\footnote{A subset of 128 centrally condensed tiles (i.e. antenna elements), designed for science cases that require large numbers of short baselines.}, a pulsar survey became computationally feasible, owing to the relatively large angular size of the tied-array beam ($\sim30^\prime$).
The Southern-Sky MWA Rapid Two-metre (SMART) pulsar survey (Bhat et al. \textit{in prep}) was conceived and began data collection in late 2018.
At present, $\sim70\%$ of the data have been collected, with the initial, first-pass (``shallow'') processing being limited to the first ten minutes (out of the full 80 minutes) of each observation.
This strategy quickly paid off with the discovery of PSR \psrnick{}, a low-luminosity, high Galactic latitude pulsar, reported in \citep{Swainston2021}.

In this paper, we report the discovery of PSR \psr{}, also detected as part of the initial search strategy.
Unlike \psrnick{}, \psr{} is sufficiently bright to be detected in single pulses, and was found to have a large nulling fraction.
Closer inspection of \psr{}'s single pulses revealed that it belongs to the class of sub-pulse drifters, a phenomenon first noted in the earliest days of pulsar research \citep{Drake1968}, and which has historically been strongly linked to magnetospheric phenomena, and thence, to the as-yet poorly understood radio emission mechanism \citep{Ruderman1975,Rankin1986,Deshpande1999}.
``Sub-pulse'' signifies discrete bursts of emission, usually much narrower than the average pulse profile, and ``drifting'' refers to the systematic way that sub-pulses arrive earlier or later in time in successive rotations.
Sub-pulse drifting is easily detected when viewing the pulses in a stack (Fig \ref{fig:pulsestacks}), in which associated sets of sub-pulses form diagonal ``drift bands'' that stretch across the on-pulse window.
The drift \textit{rate} is defined as the reciprocal of the slope of the drift bands, often expressed in units of degrees (of rotation) per pulse period ($^\circ/P$).

Many sub-pulse drifters, e.g. B0809$+$74 and B0943$+$10, are known to have very stable drift rates \citep[e.g.][]{Taylor1971,Deshpande1999}.
This stability partly motivated the original carousel model, put forward by \citet{Ruderman1975}, which associates the sub-pulses with spatially discrete bursts of electrical discharges (``sparks'') in regions of charge depletion just above the stellar surface near the magnetic poles.
According to this model, the sparks are driven by \ExB{} drift, with the observed stability of the drift rate being directly inherited from the theoretical stability of the electric and magnetic fields at the spark locations.

@@

The other type of sub-pulse drifting anomaly is where the drift rate varies relatively slowly.
This can happen in a variety of ways.
\citet{Lyne1983} linked the variability of the drift rates of B0809$+$74 and B0818$-$13 to the null sequence preceding the onset of a drift mode, showing that the drift band phase is ``remembered'' across the null.
B0826$-$34, in contrast, has a drift rate that varies slowly but pseudo randomly at all times, making the identification of discrete drift modes impossible \citep[e.g.][]{Esamdin2012}.
More recently, \citet{McSweeney2017} showed that B0031$-$07, which has long been known to exhibit three stable modes, has a measurable slow evolution of the drift rate within individual drift sequences.

These ``variations'' from stable drifting present a natural challenge to the carousel model, whose most basic formulation predicts only a single stable drifting mode, due to the expected stability of the underlying electric and magnetic fields that generate the carousel motion.
Nevertheless, many of the above effects can be successfully explained with the carousel model by invoking aliasing effects that arise due to ``beating'' between the stellar rotation rate and the carousel's rotation.
In the presence of aliasing, multiple drift modes can be explained if the number of sparks in the carousel changes, even if the carousel speed itself doesn't change.
This idea has been invoked to explain the multiple drift rates of B1918$+$19 \citep{Rankin2013}, B1819$-$22 \citep{Joshi2017}, and B0031$-$07 \citep{McSweeney2019a}.
Aliasing can also ``amplify'' the variability of the drift rate if the ratio of stellar rotation and carousel rotation is very close to the ratio of small integers.
That is, a small change in the magnetospheric conditions near the stellar surface (where the sparks reside) can appear to the observer as a relatively large change in the drift rate.
\citet{VanLeeuwen2012} elucidated the connection between the accelerating potential within the polar gap and the carousel behavior, and found that the variability of B0826$-$34's drift rate can be explained in this way.

Despite the continued success of the carousel model to explain the various observed counterexamples of stable drifting, there remain many open questions, such as what sets the number of sparks, and what drives the transition between different modes.
Exploration of these questions is hampered by the relatively small number of pulsars amenable to this kind of analysis, and is compounded by the difficulty of deriving the carousel configuration from the observed drift rate, for which there are oftentimes many degenerate solutions.
The discovery, therefore, of another bright sub-pulse drifting pulsar is a very welcome addition, particularly as this new pulsar appears to exhibit many of the rare drifting behaviors just discussed.
Section \S\ref{sec:observations} describes the observation in which the pulsar was originally found, as well a small number of archival MWA observations in which it was subsequently detected during follow up of the original candidate.
Analysis of the pulsar's nulling and drifting properties follows in Section \S\ref{sec:analysis}, which are discussed and compared to other pulsars in Section \S\ref{sec:discussion}.
We conclude with a short summary in Section \S\ref{sec:conclusions}.

@@

PSR \psr{} was discovered in one of the observations taken as part of the SMART survey (Obs ID 1226062160).
A full description of the survey parameters and observational set up will be presented in an upcoming publication \citetext{Bhat et al. in prep}.
Many details are also described in the paper reporting the first SMART pulsar discovery \citep{Swainston2021}, for which the setup was identical, but the essential details are summarised here.
The VCS delivers Nyquist-sampled dual-polarisation voltages for 128 MWA tiles ($4\times4$ cross-dipoles) at a rate of $100\,\mu$s and an instantaneous bandwidth of $30.72\,$MHz, consisting of $3072\times10\,$kHz individual channels.
Apart from the discovery observation, the other three observations presented in this work were taken as part of the follow-up timing campaign for \psrnick{} \citep{Swainston2021}.
All four observations (summarised in Table \ref{tab:observations}) were made in the frequency range $138.88$-$169.6\,$MHz.

\begin{deluxetable}{ccc}
\tabletypesize{\scriptsize}
\tablewidth{0pt}
\tablecaption{Observations of PSR \psr{} \label{tab:observations}}
\tablehead{
\colhead{MJD} & \colhead{Obs ID} & \colhead{Number of rotations}
}
\startdata
58434 & 1226062160 & 3759 \\
58991 & 1274143152 & 914  \\
59003 & 1275178816 & 914  \\
59094 & 1283104232 & 1374
\enddata
\end{deluxetable}

Each observation was individually calibrated using dedicated calibration observations of nearby bright sources and the Real Time System \citep{Mitchell2008}.
After calibration, the voltages are used to form a tied-array beam in the direction of the pulsar (the localization procedure is discussed in \S\ref{sec:localization}), using the bespoke beamforming software described in \citet{Ord2019}.
After determining the best dispersion measure (DM) and period (see \S\ref{sec:period}), de-dispsersion and folding were performed with DSPSR \citep{VanStraten2011b} and PSRCHIVE \citep{Hotan2004} to form single pulse archives, from which Stokes I pulse stacks were created.
Profiles of each detection, as well as a stacked profile incorporating the data from all four observations, is shown in Fig \ref{fig:profiles}.
Subsequent analysis of the pulse stacks is described in \S\ref{sec:analysis}.
\begin{figure}[!th]
    \centering
    \includegraphics[width=\linewidth]{profiles.png}
    \caption{Mean profiles of the four observations with detections, and the total mean profile (top panel) including all four observations. The profiles show a barely resolved double peak structure.}
    \label{fig:profiles}
\end{figure}

The brightness of \psr{}'s single pulses begs the question of how it avoided detection in previous, more sensitive pulsar surveys covering the same sky (albeit at higher frequencies).
Follow-up of the initial detection with the full 80 minutes, as well as in other serendipitous MWA observations, revealed that the pulsar has a very large nulling fraction ($\sim$ \nullfraction{}), with some nulling sequences exceeding $20\,$min (i.e. the longest archival MWA observation in which it was \textit{not} detected).
A full analysis of the nulling properties is presented in Section \S\ref{sec:nulling} below.
\psr{} may therefore be the harbinger for a population of relatively bright, intermittent pulsars which the SMART survey, with its $80$-minute dwell times, is well-suited to detect.

@@

The discovery observation (Obs ID 1226062160) was taken when the MWA was in its \textit{compact} configuration \citep[see][]{Wayth2018}, giving an effective tied-array beam size of $\sim$ 30$^\prime$ at the central frequency of $155\,$MHz.
As per the SMART survey design, these beams are tiled across the field of view, each of which is searched for periodic candidates.
Interestingly, because the compact configuration includes many redundant baselines, the tied-array beam for the compact configuration is complex, and the original detection was made in a grating lobe.
Subsequent inspection of surrounding beams revealed the tied-array beam with the brightest detection.
The ability to exploit grating lobe detections for fast-tracking candidate confirmation and localization will be discussed in the survey description paper \citetext{Bhat et al. in prep}.

>>>>

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%%
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands.
%%   tighten      : Makes baselineskip slightly smaller, only works with
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much
%%                  longer author list but may be desirable for short
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\psr}{J0026$-$1955}
\newcommand{\psrdm}{20.869}
\newcommand{\psrdmerr}{0.005}
\newcommand{\psrnick}{J0036$-$1033}
\newcommand{\dmunit}{pc$\,$cm$^{-3}$}
\newcommand{\Sd}{S$_\rm{d}$}
\newcommand{\nullfraction}{76\%}
\newcommand{\ExB}{\ensuremath{{\bf E}\times{\bf B}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Pulsar with unusual sub-pulse drifting}
\shortauthors{McSweeney et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page
%% with this command:
%% \watermark{text}
%% where "text", e.g.\ DRAFT, is the text to appear.  If the text is
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g.\ pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{figures/}}
%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

@@

\title{Independent discovery of a nulling pulsar with unusual sub-pulse
drifting properties with the Murchison Widefield Array}

@@

\author[0000-0002-0786-7307]{Samuel J. McSweeney}
\affiliation{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}
%\affiliation{Curtin Institute of Radio Astronomy, Curtin University, GPO Box U1987, Perth, 6845, WA, Australia}

\author[0000-0002-8383-5059]{N. D. Ramesh Bhat}
\affiliation{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}

\author[0000-0001-8982-1187]{Nicholas A. Swainston}
\affiliation{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}

\author[0000-0001-7801-9105]{Keegan R. Smith}
\affiliation{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}

\author[0000-0002-6631-1077]{Sanjay Kudale}
\affiliation{National Centre for Radio Astrophysics, Tata Institute of Fundamental Research, Pune 411 007, India}
%\affiliation{\dots}

\author[0000-0002-4203-2946]{Paul Hancock}
\affiliation{Curtin Institute for Computation, Curtin University, GPO Box U1987, Perth, 6845, WA, Australia}

\author[0000-0003-2519-7375]{Willem van Straten}
\affiliation{Institute for Radio Astronomy \& Space Research, Auckland University of Technology, Private Bag 92006, Auckland 1142, New Zealand}
%\affiliation{\dots}

\author[0000-0002-9618-2499]{Shi Dai}
\affiliation{Western Sydney University, Locked Bag 1797, Penrith South DC, NSW 1797, Australia}
%\affiliation{\dots}

@@

\author[0000-0002-8195-7562]{Steven J. Tingay}
\affiliation{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}

\author[0000-0003-2756-8301]{Melanie Johnston-Hollitt}
\affiliation{Curtin Institute for Computation, Curtin University, GPO Box U1987, Perth, 6845, WA, Australia}

\author[0000-0001-6295-2881]{David~L.~Kaplan}
\affiliation{Center for Gravitation, Cosmology, and Astrophysics, Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413, Milwaukee, WI 53201, USA}

\author[0000-0003-4077-6244]{Mia Walker}
\affiliation{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}

@@

We report the independent discovery of PSR \psr{} with the Murchison Widefield Array (MWA) in the ongoing Southern-sky MWA Rapid Two-meter (SMART) pulsar survey.
\psr{} has a period of $\sim1.306\,$s, a dispersion measure (DM) of $\sim\psrdm\,$\dmunit, and a nulling fraction of $\sim$ \nullfraction{}.
This pulsar highlights the advantages of the survey's long dwell times ($\sim80\,$min), which, when fully searched, will be sensitive to the expected population of similarly bright, intermittent pulsars with long nulls.
A single-pulse analysis also reveals a complex sub-pulse drifting behavior, including both rapid changes of the drift rate characteristic of mode switching pulsars, as well as a slow, consistent evolution of the drift rate within modes. In some longer drift sequences, interruptions in the otherwise smooth drift rate evolution occur preferentially at a particular phase, typically lasting a few pulses.
These properties make this pulsar an ideal test bed for prevailing models of drifting behavior such as the carousel model.

@@

Pulsars, or rapidly rotating neutron stars, are the sites of some of the highest energy physical processes in the universe, owing to the intense gravitational and magnetic fields that surround them.
Discovered first via their lighthouse-like beam of radio emission \citep{Hewish1968}, which we detect as a series of regularly spaced pulses, it was quickly realised that the physics governing the relativistic plasma that generates this beam was not well understood---a state of affairs that persists to the present day \citep{Melrose2021}.
The uniqueness of each pulsar's emission signature provides a wealth of data that can be used to test candidate emission mechanisms, and it has been proven again and again over the past several decades that in-depth studies of individual pulsars can provide valuable clues for understanding the population as a whole.
This, in part, motivates efforts to find new pulsars via large scale pulsar surveys, which continue to be conducted up to the present day.

One of the most recent such surveys is currently under way at the Murchison Widefield Array \citep[MWA;][]{Tingay2013}, a latest-generation aperture array telescope located in Western Australia.
With its Voltage Capture System \citep[VCS;][]{Tremblay2015} and offline tied-array beamforming software \citep{Ord2019,McSweeney2020}, it has proven to be a powerful instrument for investigating a variety of pulsar phenomena such as single-pulse studies \citep{McSweeney2017}, spectral analyses \citep{Meyers2017}, and studies of interstellar medium (ISM) propagation effects and related timing applications \citep{Bhat2018,Kaur2019}.
With the advent of the Phase II upgrade of the MWA \citep{Wayth2018} in its compact configuration\footnote{A subset of 128 centrally condensed tiles (i.e.\ antenna elements), designed for science cases that require large numbers of short baselines.}, a pulsar survey became computationally feasible, owing to the relatively large angular size of the tied-array beam ($\sim30^\prime$).
The Southern-Sky MWA Rapid Two-metre (SMART) pulsar survey (Bhat et al. \textit{in prep}) was conceived and began data collection in late 2018.
At present, $\sim70\%$ of the data have been collected, with the initial, first-pass (``shallow'') processing being limited to the first ten minutes (out of the full 80 minutes) of each observation.
This strategy quickly paid off with the discovery of PSR \psrnick{}, a low-luminosity, high Galactic latitude pulsar, reported in \citet{Swainston2021}.

\textbf{In this paper, we report the MWA's detection and analysis of PSR \psr{}, which emerged as a candidate as part of the same initial search strategy.}
Unlike \psrnick{}, \psr{} is sufficiently bright to be detected in single pulses, and was found to have a large nulling fraction.
Closer inspection of \psr{}'s single pulses revealed that it belongs to the class of sub-pulse drifters, a phenomenon first noted in the earliest days of pulsar research \citep{Drake1968}, and which has historically been strongly linked to magnetospheric phenomena, and thence, to the as-yet poorly understood radio emission mechanism \citep{Ruderman1975,Rankin1986,Deshpande1999}.
``Sub-pulse'' signifies discrete bursts of emission, usually much narrower than the average pulse profile, and ``drifting'' refers to the systematic way that sub-pulses arrive earlier or later in time in successive rotations.
Sub-pulse drifting is easily detected when viewing the pulses in a stack (Fig \ref{fig:pulsestacks}), in which associated sets of sub-pulses form diagonal \textit{drift bands} that stretch across the on-pulse window.
The \textit{drift rate} is defined as the reciprocal of the slope of the drift bands, often expressed in units of degrees (of rotation) per pulse period ($^\circ/P$).

Many sub-pulse drifters, e.g., B0809$+$74 and B0943$+$10, are known to have very stable drift rates \citep[e.g.][]{Taylor1971,Deshpande1999}.
This stability partly motivated the original carousel model, put forward by \citet{Ruderman1975}, which associates the sub-pulses with spatially discrete bursts of electrical discharges (``sparks'') in regions of charge depletion just above the stellar surface near the magnetic poles.
According to this model, the sparks are driven by \ExB{} drift, with the observed stability of the drift rate being directly inherited from the theoretical stability of the electric and magnetic fields at the spark locations.

@@

The other type of sub-pulse drifting anomaly is where the drift rate varies relatively slowly.
This can happen in a variety of ways.
B0809$+$74, for example, temporarily adopts a slightly lower drift rate immediately following a null, which then relaxes slowly back to the usual, stable rate after a few tens of pulses.
\citet{Lyne1983} demonstrated that this variability is related to the duration of the null sequence preceding it, and that the drift band phase is ``remembered'' across the null.
B0826$-$34, in contrast, has a drift rate that varies slowly but pseudo randomly at all times, making the identification of discrete drift modes impossible \citep[e.g.][]{Esamdin2012}.
More recently, \citet{McSweeney2017} showed that B0031$-$07, which has long been known to exhibit three stable modes, has a measurable slow evolution of the drift rate within individual drift sequences.

These variations from stable drifting present a natural challenge to the carousel model, whose most basic formulation predicts only a single stable drifting mode, due to the expected stability of the underlying electric and magnetic fields that generate the carousel motion.
Nevertheless, many of the above effects can be successfully explained with the carousel model by invoking aliasing effects that arise due to beating between the stellar rotation rate and the carousel's rotation.
In the presence of aliasing, multiple drift modes can be explained if the number of sparks in the carousel changes, even if the carousel speed itself doesn't change.
This idea has been invoked to explain the multiple drift rates of B1918$+$19 \citep{Rankin2013}, B1819$-$22 \citep{Joshi2017}, and B0031$-$07 \citep{McSweeney2019a}.
Aliasing can also ``amplify'' the variability of the drift rate if the ratio of stellar rotation and carousel rotation is very close to a ratio of small integers.
That is, a small change in the magnetospheric conditions near the stellar surface (where the sparks reside) can appear to the observer as a relatively large change in the drift rate.
\citet{VanLeeuwen2012} elucidated the connection between the accelerating potential within the polar gap and the carousel behavior, and found that the variability of B0826$-$34's drift rate can be explained in this way.

Despite the continued success of the carousel model to explain the various observed counterexamples of stable drifting, there remain many open questions, such as what sets the number of sparks, and what drives the transition between different modes.
Exploration of these questions is hampered by the relatively small number of pulsars amenable to this kind of analysis, and is compounded by the difficulty of deriving the carousel configuration from the observed drift rate, for which there are oftentimes many degenerate solutions.
The discovery, therefore, of another bright sub-pulse drifting pulsar is a very welcome addition, particularly as this new pulsar appears to exhibit many of the rare drifting behaviors just discussed.
Section \S\ref{sec:observations} describes the observation in which the pulsar was originally found and the small number of archival MWA observations in which it was subsequently detected during follow up of the original candidate.
Analysis of the pulsar's nulling and drifting properties follows in Section \S\ref{sec:analysis}, which are discussed and compared to other pulsars in Section \S\ref{sec:discussion}.
We conclude with a short summary in Section \S\ref{sec:conclusions}.

@@

PSR \psr{} was discovered in one of the observations made as part of the SMART survey (Obs ID 1226062160).
A full description of the survey parameters and observational set up will be presented in an upcoming publication \citetext{Bhat et al. in prep}.
Many details are also described in the paper reporting the first SMART pulsar discovery \citep{Swainston2021}, for which the setup was identical, but the essential details are summarised here.
The VCS delivers Nyquist-sampled dual-polarisation voltages for 128 MWA tiles ($4\times4$ cross-dipoles) at a rate of $100\,\mu$s and an instantaneous bandwidth of $30.72\,$MHz, consisting of $3072\times10\,$kHz individual channels.
Apart from the discovery observation, the other three observations presented in this work were taken as part of the follow-up timing campaign for \psrnick{} \citep{Swainston2021}.
All four observations (summarised in Table \ref{tab:observations}) were made in the frequency range $138.88$-$169.6\,$MHz.

\begin{deluxetable}{cccc}
\tabletypesize{\scriptsize}
\tablewidth{0pt}
\tablecaption{Observations of PSR \psr{} \label{tab:observations}}
\tablehead{
\colhead{MJD} & \colhead{Obs ID} & \multicolumn{2}{c}{Number of rotations} \\
& & Total & With detectable emission
}
\startdata
58434 & 1226062160 & 3759 & 788 \\
58991 & 1274143152 & 914  & 358 \\
59003 & 1275178816 & 914  & 281 \\
59094 & 1283104232 & 1374 & 545
\enddata
\end{deluxetable}

Each observation was individually calibrated using dedicated calibration observations of nearby bright sources and the Real Time System \citep{Mitchell2008}.
After calibration, the voltages are used to form a tied-array beam in the direction of the pulsar (the localization procedure is discussed in \S\ref{sec:localization}), using the bespoke beamforming software described in \citet{Ord2019}.
After determining the best dispersion measure (DM) and period (see \S\ref{sec:period}), de-dispersion and folding were performed with DSPSR \citep{VanStraten2011b} and PSRCHIVE \citep{Hotan2004} to form single pulse archives, from which Stokes I pulse stacks were created.
Profiles of each detection, as well as a stacked profile incorporating the data from all four observations, is shown in Fig \ref{fig:profiles}.
Subsequent analysis of the pulse stacks is described in \S\ref{sec:analysis}.
\begin{figure}[!th]
    \centering
    \includegraphics[width=\linewidth]{profiles.png}
    \caption{Mean profiles of the four observations with detections, and the total mean profile (top panel) including all four observations. The profiles show a barely resolved double peak structure.}
    \label{fig:profiles}
\end{figure}

\textbf{Although this pulsar emerged as a candidate in the shallow pass of SMART, it was subsequently noted to have been reported in 2018 as a candidate in the Green Bank Northern Celestial Cap (GBNCC) pulsar survey \citep{Stovall2014}.
Later re-processing of data from the 70-cm Parkes Southern Pulsar Survey  \citep{Manchester1996} also revealed a weak detection that was not previously reported.
These detections confirm the candidate as a true pulsar.}

Follow-up of the initial MWA detection with the full 80 minutes, as well as in other archival MWA observations, revealed that the pulsar has a very large nulling fraction ($\sim$ \nullfraction{})\footnote{The nulling fraction estimate comes only from analysis of the MWA data.}, with some nulling sequences potentially exceeding $20\,$min (i.e.\ the longest archival MWA observation in which it was \textit{not} detected with a periodicity search).
A full analysis of the nulling properties is presented in Section \S\ref{sec:nulling} below.
\textbf{Despite the fact that \psr{} was serendipitously discovered in the first ten minutes, it may well be} the harbinger for a population of relatively bright, intermittent pulsars which the SMART survey, with its $80$-minute dwell times, is well-suited to detect.

@@

The discovery observation (Obs ID 1226062160) was taken when the MWA was in its \textit{compact} configuration \citep[see][]{Wayth2018}, giving an effective tied-array beam size of $\sim$ 30$^\prime$ at the central frequency of $155\,$MHz.
As per the SMART survey design, these beams are tiled across the field of view, each of which is searched for periodic candidates.
The compact configuration includes many redundant baselines, and the tied-array beam for the compact configuration is complex.
Consequently, the pulsar was detected in multiple adjacent beams, including a boresight beam that yielded the highest signal-to-noise, and several nearby grating lobes.
The ability to exploit grating lobe detections for fast-tracking candidate confirmation and localization will be discussed in the survey description paper \citetext{Bhat et al. in prep}.
